---
title: "In Class Exercises - Sorting and Filtering"
author: "Sophie Holtzman"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

## Setup and Loading Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the packages we'll need
# Remember, think of this like packing for our trip
# (We don't take the whole closet, only what we need)
library(tidyverse)
library(lubridate)

```

Run this line below load the data for this assignment. It will load a dataframe called "impeach," containing U.S. House Democrats and whether they supported the first impeachment of Pres. Trump, along with when they announced such support.

```{r}

impeach <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRh8d5JaDqBtByzNw2JZF3idaACobDhhk-p7chJoktA0HawELHFOvjQOqCpzGA4MGONvPlR7GASqW-K/pub?gid=1765341510&single=true&output=csv")


```

Let's see what we have:

```{r}

impeach


```

## Exercise Questions

*For the questions below, write your R code to return the requested records in the code chunk provided.*

1)  The column "for_impeachment" indicates whether the member has publicly called for an impeachment inquiry. Filter to return only the ones where the answer is NO
-
##Used the filter command as indicated in last week's class to filter for only the NO part of the for_impeachment column. Used the same technique as the dest == SRQ command from the flight data practice.

```{r}

filter(impeach, for_impeachment == "NO")





```

2)  Filter to return only results where a member is both against impeachment, and comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column).
-
##Used the same filtering technique and added in an additional comma and and an additional column to filter. This represents anti-impeachment voters who voted Republican in 2016.

```{r}


filter(impeach, for_impeachment == "NO", p16winningparty == "R")
```

3)  Filter for only results where a member is against impeachment, comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column), and also comes from a district that Mitt Romney won in 2012 ("p12winningparty").
-
##I used the same filtering techniques, but this time I gave three columns to filter. This represents anti-impeachment voters who come from a district that voted for Trump in 2016 and Romney in 2012.

```{r}


filter(impeach, for_impeachment == "NO", p16winningparty == "R", p12winningparty =="R")


```

4)  Filter for only results from September 2019 where a member is a YES for impeachment.
-
##I used the same filtering technique, but this time I filtered using the date_month and date_year columns to bring up voters who voted in September of 2019.

```{r}


filter(impeach, for_impeachment == "YES", date_month == "9", date_year =="2019")


```

5)  Filter for only results where a member is a YES for impeachment and is from a district where Clinton won more than 70 percent of the vote in 2016 (found in column "clinton_percent").
-
##Using the same filtering technique, I added in the ">" command to filter for districts where Clinton won more than 70% of the vote.

```{r}


filter(impeach, for_impeachment == "YES", clinton_percent >= "70")


```

6)  Sort the entire dataframe based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.
-
##This time, I used the arrange command to re-arrange one of the columns in a new order of lowest to highest.

```{r}


impeach %>% 
  arrange(pct_bachelors)



```

7)  Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.
-
##In this question, I combined the filter and arrange commands to arrange for those who voted no on impeachment according to the percent of their district that has a bachelor's degree.

```{r}

filter(impeach, for_impeachment == "NO") %>%
  arrange(pct_bachelors)



```

8)  Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

Then filter those records by only those whose bachelor's percentage is below the national average (found in the "pct_bachelors_compared_to_national" column).
-
##Combining the filter and arrange commands again, I added in additional commands to filter and sort multiple columns.

```{r}


filter(impeach, for_impeachment == "NO", pct_bachelors_compared_to_national == "BELOW")  %>% arrange(pct_bachelors)



```

9)  Filter for only members from New Jersey who are NO on impeachment.
-
##I filtered using the state section and the same filtering command.

```{r}


filter(impeach, for_impeachment == "NO", state == "NJ")


```

10) Filter for those who were YES on impeachment, with a declared date prior to 2019. So only those with dates before 2019. Then sort those so that the highest Clinton vote percentages are at the top.
-
##Combining filter and arrange once again, I sorted for dates prior to 2019 and arranged percentages highest to lowest.

```{r}


filter(impeach, for_impeachment == "YES", date_year < "2019")%>%
  arrange(desc(clinton_percent))


```
